---
title: "Quarterly Report 1.2.0 Visualization"
author: "Coastal Carbon Network"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    highlight: tango
---

<style>
.vscroll-plot {
    width: 900px;
    height: 800px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

```{r setup, include=FALSE}

# this sets the working directory to start where the R project is located
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# no warnings or messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)

library(tidyverse)
library(plotly)
library(DT)
library(rnaturalearth)

```

```{r}
### Load Necessary Data ####

# library_beginnings <- tibble(
#   date = as.Date(c("2018-06-01", "2018-07-01", "2018-08-01", "2018-09-01",
#                    "2018-10-01", "2018-11-01", "2018-12-01", "2019-01-01",
#                    "2019-02-01", "2019-03-01", "2019-04-01", "2019-05-01")),
#   total_cores = c(1535, 1535, 1535, 1535, 1546, 1546, 2555, 4025, 4263, 4276,
#                   4571, 5487),
#   study_IDs = c(32, 32, 32, 32, 33, 33, 34, 188, 190, 192, 194, 254),
#   # Dated cores are inclusive of WG repo-only datasets, which can't be tapped into
#   #   from this script
#   dated_cores = c(0, 0, 0, 0, 11, 11, 22, 22, 66, 70, 133, 150)
# )
# write_csv(library_beginnings, "data/library_metrics/growth_metrics.csv")

# read in table documenting figshare pubs with IDs
versions <- read_csv("docs/synthesis_resources/synthesis_history.csv")
core_versions <- versions %>% filter(table == "cores") %>% 
  add_row(date = "3/29/24", 
          version = "1.2.0",
          filepath_id = "45374701",
          source = "figshare",
          table = "cores")

# read in current version
guess_max <- nrow(read_csv("data/CCN_synthesis/CCN_cores.csv"))

current_cores <- read_csv("data/CCN_synthesis/CCN_cores.csv", guess_max = guess_max) %>% 
  mutate(country = case_when(country == "Laos" ~ "Vietnam", 
                             country == "Russian Federation" ~ "Russia",
                             country == "Micronesia" ~ "Federated States of Micronesia",
                             T ~ country)) %>% # spot fix for Bukoski et al 2017
  # assign continent and region
  mutate(region = countrycode::countrycode(sourcevar = country,
                            origin = "country.name",
                            destination = "region"),
         continent = countrycode::countrycode(sourcevar = country,
                            origin = "country.name",
                            destination = "continent"))

# View(current_cores %>% distinct(country, region, continent))

current_version <- "1.3.0"
```

```{r}

## Create Synthesis Record (of Cores Table)

# loop through versions and harvest relevant information
if(file.exists("docs/synthesis_resources/core_synthesis_history.csv")){
  
  synthesis_history <- read_csv("docs/synthesis_resources/core_synthesis_history.csv")

  } else {
  
  synthesis_history <- data.frame()
  
  for(i in 1:nrow(core_versions)){
    
    if(core_versions$source[i] == "github"){
      temp_table <- read_csv(core_versions$filepath_id[i], guess_max = guess_max)
    } else {
      temp_table <- read_csv(paste0("https://ndownloader.figshare.com/files/", core_versions$filepath_id[i]), guess_max = guess_max)
    }
    
    # assign date and version
    temp_date <- as.Date(core_versions$date[i], format = "%m/%d/%y")
    temp_version <- core_versions$version[i]
    
    temp_df <- temp_table %>%  
      distinct(study_id, core_id) %>% 
      mutate(date = temp_date,
             version = temp_version)
    
    # compile history of synthesis in core IDs
    synthesis_history <- bind_rows(synthesis_history, temp_df)
  }
  write_csv(synthesis_history, "docs/synthesis_resources/core_synthesis_history.csv")
}

# remove any unnecessary or suspicious versions
## summary stats from this table
synthesis_history_trim <- synthesis_history %>% 
  # drop some unneccessary versions (some where the core count reduced??)
  filter(!(version %in% c("0.2.0", "1.1.0"))) %>%  # this overestimated the core count (V1.1.1 corrected this)
  filter(!(lubridate::year(as.Date(date)) %in% c(2021))) %>% 
  # add current core version info
  bind_rows(current_cores %>% 
              select(study_id, core_id) %>% 
              mutate(version = current_version, date = Sys.Date()))
  
```

## Database Snapshot

### Current Map

```{r}
# Choropleth Map of Total Cores for all Countries

country_cores <- current_cores %>% drop_na(country) %>% count(country, name = "core_count")

# world with naturalearth data
world_ne <- ne_countries(returnclass = "sf",
                     # type = "countries",
                     scale = "medium") %>% 
  rename(country = sovereignt) %>% 
  mutate(country = recode(country, "United States of America" = "United States"))
# names(world_ne)

world_country_cores <- left_join(world_ne, country_cores)

map_country_cores <- ggplot(data = world_country_cores) +
  geom_sf(color = "white", size = 0.5, aes(fill = core_count,
                                           text = paste("Country:", country, "\nTotal Cores:", core_count))) +
  coord_sf(crs = "+proj=robin") +
  theme_bw() +
  labs(fill = "Number of \nSoil Samples")

ggsave(map_country_cores, filename = "docs/quarterly_reports/v120/current_cores_map.jpg")

# print(ne_map)
ggplotly(map_country_cores + theme(legend.position = "none"), tooltip = "text") %>% 
  layout(autosize = F, width = 800, height = 500)
```

### Current Habitat Representation

```{r}

habitat_barplot <- current_cores %>% 
  # group and tally the unvegetated category
  mutate(habitat = case_when(habitat == "mudflat" ~ "unvegetated",
                          habitat == "upland" ~ "supratidal forest",
                          is.na(habitat) ~ "unknown", T ~ habitat)) %>%
  group_by(habitat) %>%
  summarise(core_count = n()) %>% 
  ungroup() %>%
  mutate(percent = 100*(core_count/sum(core_count))) %>% 
  mutate(habitat = fct_reorder(habitat, percent)) %>% 
  ggplot(aes(habitat, percent, fill = percent)) + 
  geom_col(fill = "darkgreen") + 
  coord_flip() + 
  # scale_color_brewer(palette = "BuGn") +
  xlab("Habitat Type") + ylab("Proportion of Cores (%)") +
  geom_text(aes(label = paste0("n=", core_count)), size = 3.5, hjust = -0.2) +
  # geom_text(aes(label = paste0(round(percent, 1), "%")), size = 3.5, hjust = -0.2) +
  ylim(0, 70) +
  theme_classic(base_size = 15) +
  ggtitle("Habitat Sampling Proportions")

ggsave(habitat_barplot, filename = "docs/quarterly_reports/v120/habitat_barplot.jpg", width = 6, height = 6)

  # mutate(country = recode(country, "Democratic Republic of the Congo" = "DRC")) 
  
  # ggplot(aes(core_count, country, 
  #            size = core_count,
  #            color = habitat)) + 
  # geom_point() +
  # # ggbreak::scale_x_break(c(2000, 3500)) +
  # theme_bw(base_size = 15) +
  # xlab("Number of Cores") + ylab("")

# ggplotly(p)
  # print(p)
```

## What's New

```{r}
# isolate the most recent two versions of the synthesis
previous_version <- core_versions %>% filter(version == "1.0.0") %>% 
  # filter(version != "1.1.0") %>%  # this overestimated the core count (V1.1.1 was the correction)
  # arrange(desc(version)) %>% 
  # head(1) %>% 
    pull(version)

# current <- synthesis_history_trim %>% filter(version == last_two[1]) %>% select(study_id, core_id)
previous <- synthesis_history_trim %>% filter(version == previous_version) %>% select(study_id, core_id)

# create df of new cores only
new <- anti_join(current_cores, previous)

```

**Stats:**

New studies: `r length(unique(new$study_id))`

New cores: `r nrow(new %>% distinct(study_id, core_id))`

New dated cores: `r nrow(new %>% drop_na(dates_qual_code))`


```{r fig.height=6, fig.width=10}
# stats_orig %>%
  # mutate(new_total = `Cores Available prior to project` + `New Cores (hook or publish)`,
  #        `Target Country` = recode(`Target Country`, "Federated States of Micronesia" = "Micronesia"),
  #        `Target Country` = fct_reorder(`Target Country`, new_total)) %>%
  # filter(new_total != 0) %>%

# synth_comparison <- bind_rows(left_join(previous, current_cores) %>% count(country) %>% mutate(version = previous_version),
#                         current_cores %>% count(country) %>% mutate(version = current_version))

# targets <- c("Bangladesh",
#              "Brazil",
#              "Cambodia",
#              "Cameroon",
#              "Colombia",
#              "Costa Rica",
#              "Dominican Republic",
#              "East Timor",
#              "Ecuador",
#              "Micronesia",
#              "Fiji",
#              "Ghana",
#              "India",
#              "Indonesia",
#              "Kiribati",
#              "Malaysia",
#              "Maldives",
#              "Marshall Islands",
#              "Mexico",
#              "Nauru",
#              "Palau",
#              "Papua New Guinea",
#              "Philippines",
#              "Samoa",
#              "Senegal",
#              "Solomon Islands",
#              "South Africa",
#              "Sri Lanka",
#              "Thailand",
#              "Tonga",
#              "Tuvalu",
#              "Uganda",
#              "Vanuatu",
#              "Vietnam")

synth_comparison <- full_join(left_join(previous, current_cores) %>% count(country, name = "previous_n"),
                              current_cores %>% count(country, name = "current_n")) %>% 
  mutate(previous_n = case_when(is.na(previous_n) ~ 0, T ~ previous_n), 
         change = current_n - previous_n) %>% 
  filter(change > 0) %>% 
      mutate(country = fct_reorder(country, current_n)) %>% 
  rename(Current = current_n, Previous = previous_n) %>% 
  filter(!is.na(country))
  
synth_comparison %>% select(-change) %>% 
  pivot_longer(-country, names_to = "version", values_to = "n") %>% 
  ggplot() +
  geom_point(aes(n, country, col = version), alpha = 0.5, size = 2) +
    scale_color_manual(values = c("blue", "red"), name = "Version") +
  geom_segment(data = synth_comparison, aes(y = country, yend = country, x = Previous, xend = Current), 
               colour = "black", arrow = arrow(length = unit(0.1, "cm"))) +

  ggbreak::scale_x_break(c(1800, 3800)) +
  xlab("Change in Soil Core Count") + ylab("Country") +
  theme_minimal(base_size = 12) + theme(legend.position = "bottom")

ggsave("docs/quarterly_reports/v120/country_data_progress.jpg", width = 7, height = 7)

```

```{r eval=FALSE}

bib <- read_csv("data/CCN_synthesis/CCN_study_citations.csv")

new_bib <- bib %>%  filter(study_id %in% unique(new$study_id)) %>% 
  select(-study_id) %>% distinct() %>% 
  arrange(bibliography_id)

write_csv(new_bib, "docs/quarterly_reports/v120/new_cores_bibliography.csv")

# # top contributors 
# new %>% group_by(study_id) %>% 
#   summarise(core_count = n(),
#             countries = paste(unique(country), collapse = ", "),
#             habitats = paste(unique(habitat), collapse = ", ")) %>% 
#   arrange(study_id) %>% 
#   datatable(options = list(searching = TRUE,
#                          paging = FALSE,
#                          info = FALSE,
#                          scrollY = 300,
#                          scrollX = 300,
#                          scrollCollapse = TRUE),
#           rownames = FALSE)

```

<!-- Sediment Sampling Effort -->

```{r fig.height=6, fig.width=12}
# Choropleth Map of Countries with New Cores
# resource: https://r-charts.com/spatial/maps-ggplot2/

library(rnaturalearth)

new_country <- new %>% drop_na(country) %>% count(country, name = "core_count")

# world with naturalearth data
world_ne <- ne_countries(returnclass = "sf",
                     # type = "countries",
                     scale = "medium") %>% 
  rename(country = sovereignt) %>% 
  mutate(country = recode(country, "United States of America" = "United States"))
# names(world_ne)

world_ne_cores <- left_join(world_ne, new_country) %>% 
  arrange(desc(core_count)) %>% 
  select(country, core_count, everything())

ne_map <- ggplot(data = world_ne_cores) +
  geom_sf(color = "white", width = 0.5, aes(fill = core_count,
                                           text = paste("Country:", country, "\nNew Cores:", core_count))) +
  coord_sf(crs = "+proj=robin") +
  # coord_sf(ylim = c(-60, 80), expand = FALSE) +
  # theme_void()
  theme_bw()
  # geom_point(data = data_unique, aes(x = Longitude, y = Latitude, 
                                       # color = Source), size = 3, alpha = 0.4)+
  # scale_size(range = c(2,8))+
  # theme(legend.position = "bottom")+
  # guides(col = guide_legend(ncol = 2))+
  # guides(colour = guide_legend(override.aes = list(alpha = 1)))

# Static
print(ne_map)
ggsave(ne_map, filename = "docs/quarterly_reports/v120/new_cores_map_v100_to_v120.jpg")

# Static Maps with maps library
# library(maps)
# 
# world <- map_data("world")
# sort(names(world))
# 
# ggplot(data = world) +
#   # geom_sf() +
#     geom_polygon(data = world, aes(x = long, y = lat, group = group), fill="white", color="grey")

```

```{r}
## Table to go with static plot
country_core_table <- new %>% 
  drop_na(country) %>% 
  mutate(habitat = case_when(is.na(habitat) ~ "other", T ~ habitat)) %>% 
  group_by(country) %>% 
  summarize(`total cores` = n(),
            # `dated cores` = sum(dated_core_flag),
            habitat = paste(unique(habitat), collapse = ", "),
            sources = paste(unique(study_id), collapse = ", ")) %>% 
  mutate(sources = gsub("_", " ", sources)) %>% 
  arrange(country) 
  
datatable(country_core_table, rownames = FALSE)

```

```{r eval=FALSE}
# Make it interactive
ggplotly(ne_map + theme(legend.position = "none"), tooltip = "text")
```

## Library Growth

```{r}
# plot library growth
growth_fig <- synthesis_history_trim %>% 
  count(date, version, name = "total_cores") %>% 
  
  # add a row to indicate the initial synthesis
  add_row(date = as.Date("2018-06-01"),
          version = "first",
          total_cores = 1535) %>%
  # add_row(date = Sys.Date(),
  #         version = current_version,
  #         total_cores = nrow(current_cores)) %>% 
  ggplot(aes(date, total_cores)) +
  geom_line() +
  geom_point() +
  # geom_col() +
  xlab("Date") + ylab("Soil Sample Count") +
  theme_bw()
ggsave(growth_fig, filename = "docs/quarterly_reports/v120/library_growth.jpg")

ggplotly(growth_fig) %>% layout(autosize = F, width = 800, height = 500)
```


```{r}
# derive more insight
# this is taking the approach of subsetting the current version of the synthesis
# instead of working directly with past versions 

## Approach
# The analyses should be done on the most current version of the database because 
# we revised habitats or coordinates for existing studies, it wouldn't be reflected
# work with the documentation of what cores were included in each version

habitat_change <- data.frame()
quality_change <- data.frame()
sampling_depths <- data.frame()
  
for(v in unique(synthesis_history_trim$version)){
  
  # subset synthesis history to identify studies and cores for a particular version
  temp_version <- synthesis_history_trim %>% filter(version == v) %>% 
    # make some fixes to study IDs that should align things better
    mutate(study_id = gsub("[.]", "", study_id),
           study_id = case_when(grepl(" ", study_id) ~ gsub(" ", "_", study_id), T ~ study_id))
  
  # subset present cores table to include only the cores from this particular version iteration
  # I am sure there might be a catch here
  # subset_df <- left_join(temp_version, current_cores)
  subset_df <- current_cores %>% filter(study_id %in% unique(temp_version$study_id) | core_id %in% unique(temp_version$core_id))
  
  # extract habitat information
  habitat_change <- habitat_change %>% 
    bind_rows(
      subset_df %>% count(continent, region, country, admin_division, habitat) %>% 
        mutate(version = v,
               date = unique(temp_version$date))
    )
  
  # extract quality tier information 
  subset_df_us <- subset_df %>% filter(country == "United States")
  
  quality_change <- quality_change %>%
    bind_rows(data.frame(date = unique(temp_version$date),
                         version = v,
                         total_cores = nrow(subset_df_us),
                         stock_cores = nrow(subset_df_us %>% drop_na(stocks_qual_code)),
                         dated_cores = nrow(subset_df_us %>% drop_na(dates_qual_code)),
                         elevation_cores = nrow(subset_df_us %>% drop_na(elevation_qual_code))))
  
  # harvest max sampling depths
  sampling_depths <- sampling_depths %>% 
    bind_rows(
      subset_df %>% select(continent, region, country, habitat, max_depth) %>% 
        mutate(date = unique(temp_version$date),
               version = v)
    )
}

```

### Regional Representation

```{r}
region_fig <- habitat_change %>% 
  # mutate(region = case_when(is.na(region) ~ "unknown", T ~ region)) %>% 
  drop_na(region) %>% 
  group_by(version, date, region) %>% 
  summarise(core_count = sum(n)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(date, core_count, col = region)) +
  geom_point(aes(date, core_count, col = region)) + 
  theme_bw()

ggsave(plot = region_fig, filename = "docs/quarterly_reports/v120/library_growth_region.jpg",
       height=6, width=7)
```


### Habitat

Change in Habitat Representativeness

```{r}

# "marsh" = "#607549", 
# "upland" = "#a45458",
# "mangrove" = "#c86f31ff",
# "unknown" = "#777774",
# "seagrass" = "#30b1d4",
# "unvegetated" = "#e7c687",
# "scrub/shrub" = "#749189",
# "swamp" = "#6487b8",
#  "algal mat" = "#dfc4e9"

habColors <- c("#607549", "#c86f31ff", "#30b1d4", "#a45458", "#e7c687", "#749189", "#6487b8", "#dfc4e9", "#777774" )
names(habColors) <- c("marsh", "mangrove", "seagrass", "scrub/shrub", "unvegetated", "upland", "swamp", "algal mat", "other")
colScale <- scale_colour_manual(name = "habitat", values = habColors)
# https://stackoverflow.com/questions/6919025/how-to-assign-colors-to-categorical-variables-in-ggplot2-that-have-stable-mappin

# habitat change through time
hab_fig <- habitat_change %>% 
  mutate(habitat = case_when(is.na(habitat) ~ "other", 
                             habitat %in% c("upland", "algal mat", "microbial mat") ~ "other",
                             habitat %in% c("mudflat", "sabkha") ~ "unvegetated",
                             T ~ habitat)) %>% 
  group_by(date, version, habitat) %>% 
  summarise(core_count = sum(n)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(date, core_count, col = habitat)) +
  geom_point(aes(date, core_count, col = habitat)) + 
  theme_bw() + 
  ggtitle("Data Library Growth Across Habitat Types") +
    xlab("Date") + ylab("Number of Soil Samples") +
  colScale

ggplotly(hab_fig) %>% layout(autosize = F, width = 800, height = 500)

ggsave(plot = hab_fig, filename = "docs/quarterly_reports/v120/library_growth_habitat.jpg",
       height=6, width=7)

```

```{r}
hab_region_fig <- habitat_change %>% 
    drop_na(region) %>% 
  # add holmquist synth info
    # add_row(date = as.Date("2018-06-01"),
    #         region = "North America",
    #         country = "United States",
    #         habitat = "marsh",
    #         version = "first",
    #         n = 1535) %>%
  mutate(habitat = case_when(is.na(habitat) ~ "other", 
                             habitat %in% c("upland", "algal mat", "microbial mat") ~ "other",
                             habitat %in% c("mudflat", "sabkha") ~ "unvegetated",
                             T ~ habitat)) %>% 
  group_by(date, version, region, habitat) %>% 
  summarise(core_count = sum(n)) %>% 
  ungroup() %>% 

  ggplot() +
  geom_line(aes(date, core_count, col = habitat)) +
  geom_point(aes(date, core_count, col = habitat)) + 
  facet_wrap(~region, scales = "free_y", ncol = 2) +
  xlab("Date") + ylab("Number of Soil Samples") +
  theme_bw() + colScale

ggplotly(hab_region_fig)

ggsave(plot = hab_region_fig, filename = "docs/quarterly_reports/v120/library_growth_region_habitat.jpg",
       height=7, width=7)
```

### Data Use and Completeness

Change in Data Quality Tier

```{r fig.height=6, fig.width=10}
# conver to Long format for plotting
quality_change_long <- quality_change %>% 
  mutate(other_analyses = total_cores - stock_cores) %>% 
  pivot_longer(cols = -c("date", "version"), names_to = "group", values_to = "core_n") %>% 
  filter(group != "total_cores") %>% 
  mutate(group = str_to_title(gsub("_", " ", group))) 
         # specify the order
         # group = factor(group, levels = c("Elevation Cores", "Dated Cores", "Stock Cores", "Other Analyses")))

  
# Plot 
quality_growth_fig <- ggplot(quality_change_long) + 
  # geom_area(aes(x= date, y = core_n, fill = group)) +
  geom_line(aes(date, core_n, col = group)) +
  geom_point(aes(date, core_n, col = group)) + 
  xlab("Date") + ylab("Core Count") +
  theme_bw(base_size = 15) +
  theme(legend.position = "bottom")
quality_growth_fig

ggsave(quality_growth_fig, filename = "docs/quarterly_reports/v120/library_growth_tier_US.jpg")

# ggplotly(quality_growth_fig) %>% layout(autosize = F, width = 800, height = 500)

```

### Sampling Depth

Change in Sampling Depth across habitat

```{r fig.height=6, fig.width=10}
simple_depths <- sampling_depths %>% 
  drop_na(max_depth) %>% 
  mutate(
    # three classes of sampling depth
    depth_flag = case_when(max_depth < 20  ~ "shallow (<20cm)",
                           max_depth >= 20 & max_depth <= 100 ~ "mid (20-100cm)",
                                max_depth > 100 ~ "deep (>100cm)",
                                T ~ "unknown"),
    # other categorizations
         # depth_flag = case_when(max_depth < 20 ~ "<20cm",
         #                        max_depth <= 100 ~ "20-100cm",
         #                        max_depth > 100 & max_depth <= 200 ~ "100-200cm",
         #                        max_depth > 200 & max_depth <= 300 ~ "200-300cm",
         #                        max_depth > 300 ~ ">300cm",
         #                         T ~ "unknown"),
         habitat = case_when(is.na(habitat) ~ "unknown", 
                             habitat == "mudflat" ~ "unvegetated",
                             habitat == "sabkha" ~ "unvegetated",
                              T ~ habitat)) 
  # calculate the change in composition through time
  # not as fond of that figure
  # group_by(date, version, habitat) %>% 
  # mutate(depth_flag_percent = 100*(n/sum(n)))

depth_fig <- simple_depths %>% 
    count(date, version, habitat, depth_flag) %>% 
  filter(!(habitat %in% c("unknown", "upland", "algal mat", "microbial mat"))) %>% 
  ggplot(aes(x=date, y = n, col = depth_flag)) + 
    geom_line() + geom_point() +
    facet_wrap(~habitat, scales = "free_y") +
  xlab("Date") + ylab("Core Count") +
  theme_bw()
  
ggsave(depth_fig, filename = "docs/quarterly_reports/v120/library_growth_sampling_depth.jpg")

ggplotly(depth_fig)

```

```{r}
us_sampling_depths <- simple_depths %>% 
  filter(country == "United States") %>% 
  count(date, version, habitat, depth_flag) %>% 
  filter(!(habitat %in% c("unknown", "upland", "algal mat", "microbial mat"))) %>% 
  ggplot(aes(x=date, y = n, col = depth_flag)) + 
  geom_line() + geom_point() +
  facet_wrap(~habitat, scales = "free_y") +
  xlab("Date") + ylab("Core Count") +
  theme_bw()

ggsave("docs/quarterly_reports/v120/us_sampling_depths.jpg")
```


```{r}

sampling_depths %>% 
  drop_na(max_depth) %>% 
  drop_na(region) %>%
  filter(!habitat %in% c("algal mat", "microbial mat", "upland")) %>% 
  # filter(habitat == "marsh") %>% 
  # count(date, version, continent, region, depth_flag) %>% 
  ggplot(aes(x=max_depth, y = habitat)) + 
  geom_boxplot() +
  facet_wrap(~region, scales = "free") +
  xlab("Date") + ylab("Core Count") +
  theme_bw()

```


<!-- <div class="vscroll-plot"> -->
```{r eval=FALSE}
# Kind of busy, lets leave out

# country change through time
habitat_smry <- habitat_change %>% 
  filter(!is.na(country)) %>% filter(!is.na(habitat)) %>% 
  # filter(habitat != "upland") %>% 
  # mutate(country = case_when(is.na(country) ~ "unknown", T ~ country),
         # habitat = case_when(is.na(habitat) ~ "unknown", T ~ habitat)) %>% 
  
  # filter(country == "United States") %>%
  
  # group and tally the unvegetated category
  mutate(habitat = recode(habitat, 
                          "mudflat" = "unvegetated",
                          "upland" = "other", 
                          "algal mat" = "other")) %>%
  group_by(date, version, country, habitat) %>%
  summarise(core_count = sum(n)) %>%
  ungroup()
  
  
ggplot(habitat_smry) +
  geom_line(aes(date, core_count, col = habitat)) +
  geom_point(aes(date, core_count, col = habitat)) + 
  facet_wrap(~country, scales = "free_y", ncol = 4) +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  theme_bw()

# ggplotly(country_plot)

```
<!-- </div> -->

## US State-Level Report

### Current US State Map

```{r}
# Choropleth Map of Total Cores for all Countries

us_cores <- current_cores %>% 
  filter(country == "United States") %>% 
  count(admin_division, name = "core_count") %>% 
  rename(state = admin_division)

# # world with naturalearth data
# world_ne <- ne_countries(returnclass = "sf",
#                      # type = "countries",
#                      scale = "medium") %>% 
#   rename(country = sovereignt) %>% 
#   mutate(country = recode(country, "United States of America" = "United States"))
# # names(world_ne)
# 
states <- sf::st_read("data/input_shapefiles/us_states/states_coastline_boundaries/cb_2017_us_state_500k.shp",
                  stringsAsFactors = F) %>% 
  rename(state = NAME)

state_cores <- left_join(states, us_cores)

map_state_cores <- ggplot(data = state_cores) +
  geom_sf(color = "white", size = 0.5, 
          aes(fill = core_count,
              text = paste("State:", state, "\nTotal Cores:", core_count))) +
  coord_sf(crs = "+proj=aea +ellps=WGS84 +lat_1=29.5 +lat_2=45.5 +lon_0=-96 +x_0=0 +y_0=0") +
  theme_bw() +
  labs(fill = "Number of \nSoil Samples")

ggsave(map_state_cores, filename = "docs/quarterly_reports/v120/current_cores_map.jpg")

# print(ne_map)
ggplotly(map_country_cores + theme(legend.position = "none"), tooltip = "text") %>% 
  layout(autosize = F, width = 800, height = 500)
```

### What's New

```{r}

# new_us_states <- new %>% filter(country == "United States") %>% 
#   group_by(admin_division, habitat) %>% 
#     summarize(n = n(),
#               # collapse list of unique studies for each
#               studies = paste(unique(study_id), collapse = ", "))
# 
# new_dated <- new %>% filter(!is.na(dates_qual_code)) %>% group_by(country, admin_division, habitat) %>% 
#   summarize(n = n(),
#             studies = paste(unique(study_id), collapse = ", ")) %>% 
#   arrange(admin_division)
```


**State-Level Data Progress**

```{r}
contributors_us <- new %>% 
  filter(country == "United States") %>% 
  mutate(dated_core_flag = case_when(!is.na(dates_qual_code) ~ 1, T ~ 0)) %>% 
  group_by(admin_division) %>% 
  summarize(`total cores` = n(),
            `dated cores` = sum(dated_core_flag),
            habitat = paste(unique(habitat), collapse = ", "),
            sources = paste(unique(study_id), collapse = ", ")) %>% 
  rename(state = admin_division) %>% 
  arrange(state)

datatable(contributors_us,
          options = list(searching = TRUE,
                         paging = FALSE,
                         info = FALSE,
                         scrollY = 300,
                         scrollX = 300,
                         scrollCollapse = TRUE),
          rownames = FALSE)
```

### State of the Data

**Habitat Representation**

```{r}

current_cores %>% 
  filter(country == "United States") %>% 
  group_by(habitat) %>%
  tally() %>%
  ungroup() %>%
  mutate(percent = 100*(n/sum(n))) %>% 
  mutate(habitat = fct_reorder(habitat, percent)) %>% 
  filter(!is.na(habitat)) %>% 
  ggplot(aes(habitat, percent, fill = percent)) + 
  geom_col(fill = "darkgreen") + 
  coord_flip() + 
  # scale_color_brewer(palette = "BuGn") +
  xlab("Habitat Type") + ylab("Proportion of Cores (%)") +
  geom_text(aes(label = paste0("n=", n)), size = 3.5, hjust = -0.2) +
  # geom_text(aes(label = paste0(round(percent, 1), "%")), size = 3.5, hjust = -0.2) +
  ylim(0, 80) +
  # theme_classic() +
  theme_classic(base_size = 15) +
  ggtitle("Habitat Sampling Proportions (US)")

ggsave("docs/quarterly_reports/v120/US_habitat_pct_barplot.jpg", width = 6, height = 6)
```

### Blue Carbon Report Card

**COMING SOON**

## Change in US States

```{r}
growth_fig_us <- habitat_change %>% 
  filter(country == "United States") %>% 
    mutate(habitat = case_when(is.na(habitat) ~ "other", 
                             habitat %in% c("upland", "algal mat", "microbial mat") ~ "other",
                             habitat %in% c("mudflat", "sabkha") ~ "unvegetated",
                             T ~ habitat)) %>% 
  group_by(version, date) %>% 
  summarise(core_count = sum(n)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(date, core_count)) +
  geom_point(aes(date, core_count)) + 
  xlab("Date") + ylab("Soil Sample Count") +
  theme_bw(base_size = 15)

ggsave(growth_fig_us, filename = "docs/quarterly_reports/v120/library_growth_us.jpg")


```

```{r}
# this should be normalized by the area of these habitats

habitat_change %>% 
  filter(country == "United States") %>% 
    mutate(habitat = case_when(is.na(habitat) ~ "other", 
                             habitat %in% c("upland", "algal mat", "microbial mat") ~ "other",
                             habitat %in% c("mudflat", "sabkha") ~ "unvegetated",
                             T ~ habitat)) %>% 
  group_by(version, date, habitat) %>% 
  summarise(core_count = sum(n)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(date, core_count, col = habitat)) +
  geom_point(aes(date, core_count, col = habitat)) + 
  xlab("Date") + ylab("Soil Sample Count") +
  theme_bw()
```

