---
title: "CCN Synthesis Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    highlight: tango
---
  
<style>
.vscroll-plot {
    width: 900px;
    height: 800px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>
  
```{r setup, include=FALSE}
# this sets the working directory to start where the R project is located
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# no warnings or messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# define custom options for data tables
custom_opts <- list(searching = TRUE,
                    paging = FALSE,
                    info = FALSE,
                    scrollY = 300,
                    scrollX = 300,
                    scrollCollapse = TRUE,
                    fixedColumns = list(leftColumns = 2))

# load libraries
library(leaflet)
library(DT)
```
  
## Introduction 
The CCN clearinghouse provides a synthesized and curated set of coastal carbon data that promotes the sharing of open data, novel analyses, and collaboration opportunities. The synthesis is created by curating and merging publicly available data through an automated script. This report documents the success or failure of the automated synthesis process, documents changes, conducts QA/QC tests, and provides visualizations and summaries of data within the clearinghouse. 

***

**Report date: `r formated_date`**

**Synthesis status: `r if(join_status) {"Successful"} else "Failed"`**

**Contact: Jaxine Wolfe (wolfejax@si.edu)**

***

**Database Overview**

Currently, this synthesis contains `r length(unique(ccrcn_synthesis$depthseries$study_id))` studies with soil carbon associated observations across `r length(unique(ccrcn_synthesis$depthseries$site_id))` sites and `r length(unique(ccrcn_synthesis$depthseries$core_id))` cores from around the world. Observations represent a `r max(as.numeric(ccrcn_synthesis$cores$year), na.rm = T) - min(as.numeric(ccrcn_synthesis$cores$year), na.rm = T)` year timespan, from `r min(as.numeric(ccrcn_synthesis$cores$year), na.rm = T)` to `r max(as.numeric(ccrcn_synthesis$cores$year), na.rm = T)`.

```{r, echo=FALSE}

library(leaflet)
library(viridis)

# filter core table as needed for investigation
map_cores <- ccrcn_synthesis$cores %>%
  mutate(habitat = case_when(habitat == "mudflat" ~ "unvegetated", 
                             TRUE ~ habitat)) %>% 
  # current habitats: "algal mat", "mangrove", "marsh", "mudflat", "scrub shrub", "seagrass", "swamp", "unvegetated", "upland"
  # filter(habitat == "mangrove") %>% 
  drop_na(longitude) 

habs <- unique(map_cores$habitat)

# pal <- colorFactor(palette = 'Dark2', domain = map_cores$habitat)
pal <- colorFactor(viridis_pal(option = "H")(length(habs)), domain = habs)

leaflet(map_cores, width = "100%") %>%
  addTiles() %>%
  addCircleMarkers(lng = ~as.numeric(longitude), lat = ~as.numeric(latitude), radius = 1,
                   label = ~paste0(study_id, "; ", habitat), color = ~pal(habitat)) %>%
  addLegend(pal = pal, values = ~habitat)

```

## QA/QC

### Test Results Overview 

```{r, echo=FALSE}

table_length <- nrow(qa_results)

# data table of results 
datatable(qa_results, options = list(pageLength = table_length,
                                     autoWidth = TRUE,
                                     columnDefs = list(list(width = "47%", targets = c(1,2))),
                                     searching = FALSE,
                                     paging = FALSE))

```

### Relational Tables

Cases where a core ID is present in the core-level table but not the depthseries.

```{r, echo=FALSE}
core_results <- anti_join(ccrcn_synthesis$cores, ccrcn_synthesis$depthseries, by="core_id") %>% 
  distinct(study_id, site_id, core_id)

datatable(core_results,
          options = list(searching = TRUE,
                         paging = FALSE,
                         info = FALSE,
                         scrollY = 300,
                         scrollX = 300,
                         scrollCollapse = TRUE),
          rownames = FALSE)

```

Cases where a core ID is present in the depthseries table but not the core-level table.

```{r, echo=FALSE}

depthseries_results <- anti_join(ccrcn_synthesis$depthseries, ccrcn_synthesis$cores, by="core_id") %>%
  distinct(study_id, site_id, core_id)

datatable(depthseries_results,
          options = list(searching = TRUE,
                         paging = FALSE,
                         info = FALSE,
                         scrollY = 300,
                         scrollX = 300,
                         scrollCollapse = TRUE),
          rownames = FALSE)

```

### Uncontrolled Attributes

```{r, echo=FALSE}

# database_structure <- read_csv("docs/ccrcn_database_structure.csv", col_types = cols())

# Extract column names for that table and subset to all columns that aren't in the controlled or uncontrolled attribute tables
core_cols <- filter(database_structure, table == "cores")$attribute_name
ds_cols <- filter(database_structure, table == "depthseries")$attribute_name
methods_cols <- filter(database_structure, table == "methods")$attribute_name

join_uncontrolled_attributes <- data.frame()

if(!all(names(ccrcn_synthesis$methods) %in% methods_cols)){
  methods_uncontrolled <- ccrcn_synthesis$methods %>%
  select(study_id | !any_of(methods_cols)) %>% 
  pivot_longer(-study_id, names_to = "attribute_name", values_to = "value") %>% 
  drop_na(value) %>% distinct() %>% 
  mutate(table = "methods") %>% select(table, everything())
  
  join_uncontrolled_attributes <- bind_rows(join_uncontrolled_attributes, methods_uncontrolled)
}

if(!all(names(ccrcn_synthesis$cores) %in% core_cols)){
cores_uncontrolled <- ccrcn_synthesis$cores %>%
  select(study_id | !any_of(core_cols)) %>% 
  pivot_longer(-study_id, names_to = "attribute_name", values_to = "value") %>% 
  drop_na(value) %>% distinct() %>% 
  mutate(table = "cores") %>% select(table, everything())

  join_uncontrolled_attributes <- bind_rows(join_uncontrolled_attributes, cores_uncontrolled)

}

if(!all(names(ccrcn_synthesis$depthseries) %in% ds_cols)){
ds_uncontrolled <- ccrcn_synthesis$depthseries %>%
  select(study_id | !any_of(ds_cols)) %>% 
  pivot_longer(-study_id, names_to = "attribute_name", values_to = "value") %>% 
  drop_na(value) %>% distinct() %>% 
  mutate(table = "depthseries") %>% select(table, everything())

  join_uncontrolled_attributes <- bind_rows(join_uncontrolled_attributes, ds_uncontrolled)
}

if(nrow(join_uncontrolled_attributes) > 0){
  join_uncontrolled_attributes <- join_uncontrolled_attributes %>% 
  mutate(value = case_when(grepl("[...]", attribute_name) ~ "row number index", # highlight row number index columns with a flag
                           grepl("([0-9]+).*$", value) ~ "numeric value", # replace numeric values with a flag
                           T ~ value)) %>% 
  distinct() %>% arrange(study_id)
}
```

The studies with uncontrolled attributes in essential tables are the following: 

`r unique(join_uncontrolled_attributes$study_id)`

This table provides a more detailed breakdown of these attributes and the values associated with them:

```{r, echo=FALSE}

datatable(join_uncontrolled_attributes,
          options = list(searching = TRUE,
                         paging = FALSE,
                         info = FALSE,
                         scrollY = 300,
                         scrollX = 300,
                         scrollCollapse = TRUE),
          rownames = FALSE)
```

### Uncontrolled Variables

```{r}

uc_vars <- str_split(toString(
  qa_results %>% 
    filter(grepl("Undefined variables", result)) %>% 
    mutate(result = gsub("Undefined variables: ", "", result)) %>% 
    pull(result)), 
  pattern = ", ")[[1]]

# use this table as a hack while I work on a better solution
uncontrolled <- read_csv("docs/resolve_uncontrolled.csv") %>% 
  filter(variable_name %in% uc_vars) %>% 
  filter(attribute_name != "coring_method") %>% 
  select(-c(action, definition, notes, uncontrolled_type)) %>% 
  bind_rows(ccrcn_synthesis$methods %>% 
              filter(coring_method %in% uc_vars) %>% select(study_id, coring_method) %>% 
              rename(variable_name = coring_method) %>% 
              mutate(table_name = "methods", attribute_name = "coring_method")) %>% 
  distinct() %>% arrange(study_id)
# theres a few that aren't included, but it's better than nothin

datatable(uncontrolled,
                    options = list(searching = TRUE,
                         paging = FALSE,
                         info = FALSE,
                         scrollY = 300,
                         scrollX = 300,
                         scrollCollapse = TRUE),
          rownames = FALSE)
```


***

### Numeric Attributes

**Summary Statistics:**
```{r, echo=FALSE}

# data table of results 
datatable(qa_numeric_results,
          extensions = "FixedColumns",
          options = custom_opts,
          rownames = FALSE)

```

### Bibliography

```{r, echo=FALSE}
no_citations <- ccrcn_synthesis$cores %>%
  filter(!(study_id %in% unique(ccrcn_synthesis$study_citations$study_id)))

# If data was removed, print a warning statement
if (nrow(no_citations) > 0) {
  print("The following studies do not have citation information present.")
  print(unique(no_citations$study_id))
}
```


```{r, echo=FALSE}

bib_file_redundancy_checks <- ccrcn_synthesis$study_citations %>%
  select(-c(study_id, publication_type, keywords,
  #           # issn, `article-number`,
            abstract)) %>%
  drop_na(bibliography_id) %>% distinct_all() %>% 

  group_by(bibliography_id) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)

if(nrow(bib_file_redundancy_checks) != 0){
  print("There are duplicate bibliography IDs. Please see the following table:")
  print(bib_file_redundancy_checks)
}

```

***

## Stats and Visualization 

### Habitats

```{r, echo=FALSE, eval=FALSE}
habitat_smry <- map_cores %>% 
  count(habitat) %>% 
  mutate(pct = round(100*(n/sum(n)), 2),
         habitat = ifelse(is.na(habitat), "unknown", habitat)) %>% 
  arrange(desc(pct))

knitr::kable(habitat_smry)
```

```{r, echo=FALSE}
ccrcn_synthesis$cores %>% 
  mutate(habitat = ifelse(is.na(habitat), "unknown", habitat)) %>%
  group_by(habitat) %>%
  tally() %>%
  ungroup() %>%
  mutate(percent = 100*(n/sum(n))) %>% 
  mutate(habitat = fct_reorder(habitat, percent)) %>% 
  # filter(!is.na(habitat)) %>% 
  ggplot(aes(habitat, percent, fill = percent)) + 
  geom_col(fill = "darkgreen") + 
  # scale_color_brewer(palette = "BuGn") +
  xlab("Habitat Type") + ylab("Proportion of Samples (%)") +
    geom_text(aes(label = n), size = 2.75, hjust = -0.2) +

  # geom_text(aes(label = paste0(round(percent, 1), "%")), size = 2.75, hjust = -0.2) +
  # ylim(0, 50) +
  coord_flip() +
  theme_bw() # base_size = 17
```

The following table shows which cores are currently not assigned any habitat:

```{r, echo=FALSE}

# habitat
missing_habitat <- ccrcn_synthesis$cores %>% 
  filter(is.na(habitat)) %>% 
  select_if(function(x) {!all(is.na(x))})

datatable(missing_habitat,
          # extensions = "FixedColumns",
          options = list(searching = TRUE,
                         paging = FALSE,
                         info = FALSE,
                         scrollY = 300,
                         scrollX = 300,
                         scrollCollapse = TRUE),
          rownames = FALSE)
```

### Geography

The following table shows which cores are currently not assigned any country. This may be because the latitude and longitude are undefined.

```{r, echo=FALSE}

# country
missing_country <- ccrcn_synthesis$cores %>% 
    filter(is.na(country)) %>%
  select(study_id, site_id, core_id, latitude, longitude) %>% 
  mutate(latitude = ifelse(is.na(latitude), "unknown", latitude), 
         longitude = ifelse(is.na(longitude), "unknown", longitude))

datatable(missing_country,
          # extensions = "FixedColumns",
          options = list(searching = TRUE,
                         paging = FALSE,
                         info = FALSE,
                         scrollY = 300,
                         scrollX = 300,
                         scrollCollapse = TRUE),
          rownames = FALSE)

```

***

### Depthseries

**Depth Interval Checks**

```{r, echo=FALSE, fig.width=12, fig.height=12}
ggplot(ccrcn_synthesis$cores, aes(max_depth, col = habitat)) + geom_density() + geom_rug() +
  facet_wrap(~habitat, scales = "free", ncol = 3) +
  theme(legend.position = "none") +
  ggtitle("Distribution of Max Sampling Depths by Habitat")
```

**The following studies have cores with a max depth of NA or 0: `r ccrcn_synthesis$cores %>% filter(is.na(max_depth) | max_depth == 0) %>% distinct(study_id) %>%  pull(study_id)`**

```{r, echo=FALSE}
# Isolate instances of multiple observations for the same interval within a unique core
dup_interval <- ccrcn_synthesis$depthseries %>% 
  add_count(study_id, site_id, core_id, method_id, depth_min, depth_max) %>% filter(n > 1) %>% 
  distinct(study_id, site_id, core_id, method_id, n)

# Identify single interval samples in the depthseries and classify them
one_interval <- ccrcn_synthesis$depthseries %>% 
  add_count(core_id) %>% filter(n == 1) %>% 
  mutate(across(c(depth_min, depth_max), as.numeric)) %>% 
  mutate(sample_type = case_when(depth_min == 0 & depth_max < 20 ~ "surface sample",
                                 representative_depth_min == 0 & depth_max < 20 ~ "surface sample",
                                 depth_min != 0 ~ "subsurface sample",
                                 T ~ "single large interval")) %>% 
  select(study_id, site_id, core_id, method_id, depth_min, depth_max, sample_type)

# ggplot(one_interval) + 
#   geom_segment(aes(x = depth_min, y = core_id, 
#                    xend = depth_max, yend = core_id))

# check that the depth max is always greater than the depth min
depth_check <- ccrcn_synthesis$depthseries %>% 
  mutate(depth_diff = as.numeric(depth_max) - as.numeric(depth_min)) %>% 
  filter(depth_diff < 0) %>% 
  select(study_id, site_id, core_id, depth_min, depth_max)

```

Instances of multiple observations for the same interval within a unique core (and unique method ID) were found in the following `r length(unique(dup_interval$study_id))` studies: 

`r sort(unique(dup_interval$study_id))`

There are `r nrow(one_interval %>% filter(sample_type == "surface sample"))` single-interval surface samples (max depth < 20cm) and `r nrow(one_interval %>% filter(sample_type == "subsurface sample"))` single-interval subsurface samples. There are also `r nrow(one_interval %>% filter(sample_type == "single large interval"))` large single-interval samples (ex. one representative (?) sample for 0-100cm depth).

The following `r length(unique(depth_check$study_id))` studies had at least one depth interval where the depth max was less than the depth min: `r sort(unique(depth_check$study_id))`

**Numeric Data Distributions**

<div class="vscroll-plot">

```{r, echo=FALSE, warning=FALSE, fig.width=12, fig.height=25}

ds_numcols <- database_structure %>% filter(table == "depthseries" & data_type == "numeric") %>% 
  distinct(attribute_name) %>% pull(attribute_name)

ccrcn_synthesis$depthseries %>%
  select(any_of(ds_numcols)) %>%
  mutate(across(everything(), as.numeric)) %>%
  gather() %>%
  ggplot(aes(value)) +
  facet_wrap(~key, scales = "free", ncol = 4) +
  geom_density() + geom_rug()
# 
# # density plot
# ccrcn_synthesis$depthseries %>%
#   select(any_of(ds_numcols)) %>%
#   mutate(across(everything(), as.numeric)) %>% 
#   gather() %>%
#   ggplot(aes(value)) +
#   facet_wrap(~key, scales = "free") +
#   # geom_histogram()
#   geom_density()
#   # geom_rug()
```
</div>

```{r, echo=FALSE, fig.width=12, fig.height=12, warning=FALSE}

# things to plot:
# density of values (ex. DBD)
# downcore values?

# table_merge <- ccrcn_synthesis$depthseries %>% 
#   left_join(ccrcn_synthesis$cores %>% select(contains("_id"), habitat))
# 
# table_merge %>% 
#   drop_na(fraction_carbon, fraction_organic_matter) %>% 
#   ggplot(aes(fraction_organic_matter, fraction_carbon, col = habitat)) +
#   geom_point(alpha = 0.5, size = 1.5) +
#   theme_bw(base_size = 15) + 
#   # facet_wrap(~habitat, scales = "free", dir = "v") +
#   theme(legend.position = "bottom")

# study-level summary table of

```

**Relationship between organic matter (LOI) and carbon content**

<div class="vscroll-plot">
```{r OM-C, echo=FALSE, fig.width=12, fig.height=30, warning=FALSE}

ccrcn_synthesis$depthseries %>% 
  drop_na(fraction_organic_matter, fraction_carbon) %>% 
  # add_count(study_id) %>% filter(n > 10) %>%
  # these studies havce modeled values
  # filter(study_id %in% c("Drexler_et_al_2009", "van_Ardenne_et_al_2018", "Ceron-Breton_et_al_2011",
                         # "Keshta_et_al_2020", "Rodriguez_et_al_2022", "Hamzeh_and_lahijani_2022")) %>%
  ggplot(aes(as.numeric(fraction_organic_matter), as.numeric(fraction_carbon))) + 
  geom_point(alpha = 0.5, col = "darkgreen") + 
  xlab("Fraction Organic Matter") + ylab("Fraction Carbon") +
  facet_wrap(~study_id, scales = "free", ncol = 4)

```
</div>

***

## Change Log

**COMING SOON**

The following table summarizes which entries across the range of synthesis tables changed compared to the previous synthesis. Study IDs associated with the "forward" change type represent data that was not present in the previous synthesis, while "backward" indicates that study ID is no longer in the given table. If a study ID in a table is associated with both backward and forward change types, it is likely that one or more values were altered for that row in the new synthesis. 

```{r, echo=FALSE}

  # # data table of change log results 
  # datatable(change_log_results)
  # 
  # if(nrow(change_log_errors) > 0){
  #   datatable(change_log_errors)
  #   }

# document new data added & from which studies
# new species?
# document data gaps? => studies that need methods?
```

***

## Miscellaneous
The following files were found in the derivative folders. They either do not match the approved filename structure for csvs or are another filetype

#### Unknown CSVs

```{r, echo=FALSE}

if(!is.null(file_paths$unknown_csv)){
  print(unique(file_paths$unknown_csv))
}else {
  "No unknown csv files found in derivative folders"
}

```

#### Non .csv or .bib filetypes

```{r, echo=FALSE}

if(!is.null(file_paths$unknown_filetypes)){
  print(file_paths$unknown_filetypes)
} else {
  "No non-csv or -bib files found in derivative folders"
}
```

#### Warnings Encountered

```{r, echo=FALSE}
print(warning_summary)
```
