---
title: "QAQC Report"
author: "Coastal Carbon Research Coordination Network"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    highlight: tango
---

```{r setup, include=FALSE}
# this sets the working directory to start where the R project is located
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# no warnings or messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)

# load libraries
library(leaflet)
library(DT)

# source QA functions
# source("scripts/1_data_formatting/qa_functions.R")
```

### Attribute tests

```{r}
coltest <- testTableCols(table_names) %>% mutate(attribute_test = "validity")
reqtest <- testRequired(table_names) %>% mutate(attribute_test = "required")
contest <- testConditional(table_names) %>% mutate(attribute_test = "conditional")

attributetest <- bind_rows(coltest, reqtest) %>% 
  bind_rows(contest) %>% 
  arrange(table) %>% 
  select(table, attribute_test, result)
```

Validity: whether or not a column name is present in the CCN guidance
Required:: if all required attributes are present
Conditional: if all conditional attributes are present 

### Variable test
```{r}
vartest <- testTableVars(table_names)

if(vartest == "Passed"){
  "Passed! All variable names match CCRCN standards."
} else {
  "View resulting invalid variable names:"
  knitr::kable(vartest)
}

# <a href="#" class="btn btn-success">Success</a>

```

### Uniqueness

Test for non-unique core IDs

```{r}
unicores <- testUniqueCores(cores)

if(unicores == "Passed"){
  "Passed! All core IDs are unique."
} else {
  "Check on the following cores IDs:"
  knitr::kable(unicores)
}
```

Test for non-unique core locations 

```{r}
unicoords <- testUniqueCoords(cores)

if(unicoords == "Passed"){
  "Passed! No duplicate core locations detected."
} else {
  "View resulting list of coordinates with more than one associated core:"
  datatable(unicoords,
            options = list(searching = FALSE,
                           paging = FALSE,
                           info = FALSE,
                           scrollY = 300,
                           scrollX = 300,
                           scrollCollapse = TRUE),
            rownames = FALSE) 
}
```

### Relational structure

Site IDs in core and depthseries tables: `r testIDs(cores, depthseries, by = "site")`
```{r}
testIDs(cores, depthseries, by = "site") # revise output 
```

Core IDs in the core and depthseries tables: `r testIDs(cores, depthseries, by = "core")`
```{r}

testIDs(cores, depthseries, by = "core") # revise output
```

### Numerical values

Check whether fractions are expressed as fractions (and not percentages):

```{r}
fractiontest <- fractionNotPercent(depthseries)

if(fractiontest == "Passed"){
  "Passed! All fractions are expressed as fractions."
} else if(fractiontest == "No fraction columns"){
  "Passed! No fraction columns present in the dataset."
} else {
  "At least one of the attributes intended to be expressed as a fraction is expressed as a percent (values > 1). Please review all fraction columns:"
  fractiontest
}
```

Numerical ranges

```{r}
numtest_cores <- testNumericCols(cores) %>% mutate(table = "cores")
numtest_ds <- testNumericCols(depthseries) %>% mutate(table = "depthseries")
skim_numerical <- bind_rows(numtest_cores, numtest_ds) %>% select(table, everything())

datatable(skim_numerical,
          # extensions = "FixedColumns",
          options = list(searching = TRUE,
                         paging = FALSE,
                         info = FALSE,
                         scrollY = 300,
                         scrollX = 300,
                         scrollCollapse = TRUE),
                         # fixedColumns = list(leftColumns = 2),
          rownames = FALSE)
```

```{r}
if(any(grepl("activity", names(depthseries)))){
  # unitcols <- names(depthseries)[grepl("unit", names(depthseries))]
  dating_units <- depthseries %>% select(contains("unit")) %>% 
    fill(everything(), .direction = "downup") %>% distinct() %>% 
  
  print("Study-specific dating units:")
  knitr::kable(dating_units)
}
```

